Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2017-11-09 docs/thesis/chapters/1_introduction/introduction.tex
chapter 2 ok

Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}. Today organizations and enterprises are adopting the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, the corporate mobile client is usually light-weighted and generally less protected.

@@

The high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms. Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection. Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms. Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high thoughput data analysis.

@@

In some applications, the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks. Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}. Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}. Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results. However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.

@@

This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to corporate mobile apps, which are the expired user misusing password and the intruder attack. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of this proposed solution is to provide a maximum defense at the minimal resource cost.

Considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, and also considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, we also propose a tensor-based dictionary learning method for fraud detecton from unbalanced data, in order to apply the tensor deconposition for dictionary learning methods in order to evidentiate the discriminative sensing of a fraud detection dataset. Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions. We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.

@@

The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.

architecture and performance evaluation for mobile behavioral analysis

an adaptive fraud detecion technique for imbalanced data

applied to security analysis, such as network attack detection, malicious behavios in mobile apps and fraud detection in mobile payment transacions

@@

This thesis is organized as follows. In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis. Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application. Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions. Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work. Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

>>>>

Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized, leading to obscure and incomplete security paradigms and solutions \cite{galibus2017offline}.

Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}. Today organizations and enterprises are adopting the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms, being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, the corporate mobile client is usually light-weighted and generally less protected.

@@

The high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms. Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.

Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms. Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high thoughput data analysis.

@@

In some applications, the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks. Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}. Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}. Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results. Additionally, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.

@@

This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to offline corporate mobile apps, which are the expired user misusing password and the intruder attack. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of this proposed solution is to provide a maximum defense at the minimal resource cost.

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset. Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions. We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.

@@

We analyse problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques. The results of the work presented in this thesis provide the following contributions:

\begin{enumerate}
	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset;
	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
	\item We propose an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discuss the processing time of the proposed framework for mobile devices;
	\item We propose a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
\end{enumerate}

@@

This thesis is organized as follows. In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis. Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application. Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions. Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work. Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2017-11-09 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
chapter 2 ok

% In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.

@@

We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic. This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.

@@

\begin{algorithm}[h!]
	\caption{Detection of Time Frames Under Attack}
  	\label{alg:2_alg01}
	% \begin{algorithmic}[1]
	% 	\show\LOOP
	%     \ALOOP {$f = 1$ until $f == \hat{d}$} 
	%     	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
 %    	    	    \ALOOP {$i = 1$ until $i == Q$} 
	% 			\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
	% 			    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
	% 			\ENDIF
 %        		\ENDALOOP
	%     		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
 %   		\ENDALOOP
	% \end{algorithmic}
\end{algorithm}

After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked. To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.

@@

This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$. Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.

@@

This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

@@

This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

>>>>

In this work, scalars are denoted by italic letters ($a, b, A, B, \alpha, \beta$), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.

@@

We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic. This work shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.

@@

\begin{algorithm}[h!]
	\caption{Detection of Time Frames Under Attack}
  	\label{alg:2_alg01}
	\begin{algorithmic}[1]
		\For{$f = 1$ until $f == \hat{d}$} 
    	    \State $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
	    	\For{$i = 1$ until $i == Q$} 
				\If {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
			    	\State $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
				\EndIf
    		\EndFor
    		\State $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
    	\EndFor
	\end{algorithmic}
\end{algorithm}

After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected Attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked. To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.

@@

This work proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$. Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.

@@

This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

@@

This work evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
